% !TeX program = lualatex

\documentclass{article}
\usepackage{tikz}
\usepackage[utf8]{inputenc}
\setlength{\parindent}{0em} %indents to paragraphs
\setlength{\parskip}{1em} %lineskips after paragraph breaks
\usepackage[margin=1.0in]{geometry} %margins
\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{tikz-feynman}
\usepackage{amsmath}
\usepackage{physics}
\usetikzlibrary{calc}
\usepackage{feynmp}
\usepackage{feynmp-auto}
\usepackage{breqn}
\usepackage{graphicx}
\newenvironment{amatrix}[1]{%
	\left(\begin{array}{@{}*{#1}{c}|c@{}}
	}{%
	\end{array}\right)
}
\makeatletter
\let\@@span\span
\def\sp@n{\@@span\omit\advance\@multicnt\m@ne}
\makeatother

\renewcommand{\span}{...}


\title{AST121: Assignment 2}
\author{A. Lehmann}
\date{March 2023}

\begin{document}
	\maketitle
	\textit{Assignment 2 iterates upon the concepts introduced in Assignment 1. Hence it is advised to read Assignment 1 before Assignment 2. }
	\tableofcontents
	\newpage 
	\section{Calculation of distance, size, and Hubble Constant for two galaxy clusters}
	\textbar{We observe two arbitrary galaxy clusters, and we find Cluster A is situated at a distance of 50 Mpc, as extrapolated from using Cepheid variables. In addition, we have observed that the galaxies in Cluster B are roughly 64 times dimmer than those in Cluster B. We know the distance of Cluster A. Now how does one uncover how far Cluster B lies from us?}
	\subsection{The Inverse Square Law}
	
	\textbar{The inverse square law describes the relationship between energy density per unit of time and the distance from which we observe a light source, which is almost ubiquitous in many equations describing the workings of the cosmos, such as Newtons universal law of gravitation, which states that the strength of a gravitational field decreases with the square of the distance.}
	\begin{gather*}
	F_g = \frac{GMm}{r^2} \implies F_g \propto \frac{1}{r^2}
	\end{gather*}

	\textbar{One can come about the inverse square law when pondering the geometry of light dispersing out in a three-dimensional space. When light radiates from a point source, it spreads out evenly in all directions, forming a sphere of light. As the sphere expands, the same amount of light is distributed over a larger and larger surface area. Therefore, the light's intensity decreases, with the square of the area, as it travels further away from its light source.}
	\begin{center}
		\textit{Surface Area of Sphere = $4\pi r^2, $ where $r$ is the radius}
	\end{center}

\textbar{Now, we can utilize Maxwell's first equation in integral form to derive the inverse square law.}
\begin{center}
\textit{Maxwell's first equation, (refered to as Gauss's law):}
$\boxed{\int E \cdot d\vec{a}  = \frac{Q_{encl}}{\epsilon_0}}
$

\end{center} 
\textit{We may rewrite for the derivation of inverse square law: }
\begin{gather}
\iint_{\mathcal{S}_c} E \cdot d\vec{a} = 4\pi \iiint_V \rho dV 
\end{gather}
\begin{gather}
E(r) \iint_{\mathcal{S}_c} d\vec{a} = 4\pi r^2 E(r)
\end{gather}
\begin{gather}
4\pi r^2 E(r) = 4\pi e
\end{gather}
\begin{center}
	$\boxed{E(r) = \frac{e}{r^2}}$
	
\end{center}


%%%ADD MORE EXPLAINTIONS TO HOW I DERIVIED THIS BEAST%%%

\textbar{From our understanding of inverse square law, we can conclude:}
	\begin{gather*}
	I = \frac{L}{4\pi r^2} \implies I \propto \frac{1}{r^2}
	\end{gather*}
\textbar{Now we can solve for the distance to Cluster B. We know the distance to Cluster A is 50 Mpc, and the galaxies in Cluster B are 64 times fainter than A. With inverse square law, we may write this as:}
\begin{gather*}
	I \propto \frac{1}{r^2} \implies \Delta I = \frac{d_{B}^2}{d_{A}^2}
\end{gather*}
\begin{gather*}
\Delta I \times d^2_A = d^2_B
\end{gather*}
\begin{center}
$\boxed{ d_B = \sqrt{\Delta I \times d^2_A} = \sqrt{64 \times 50^2Mpc} = 400Mpc}
$
\end{center}
\textbar{Using the inverse square law, we deduced that Cluster B is approximately 400 Mpc away from us.}	

\subsection{Cosmological dimming and its influence on the apparent size of distant objects}

\textbar{If one were to probe the sky for these two cosmic objects, one would notice an apparent delta in brightness and size when comparing Cluster A and B. Cluster B would appear noticeably smaller and fainter. We may explain by the inverse square law, for when the light from the galaxies in Cluster has travelled a greater distance throughout space, causing it to become dimmer. As a result, Cluster B takes up a smaller subset of the sky compared to Cluster A. However, if we consider the luminosity of these two clusters, we observe them to be approximately the same.
	
The connection between the size of an object and its luminosity can vary depending on the class of cosmic object we are studying. Commonly, larger objects tend to be more luminous than smaller objects of the same kind. The relation exists because larger objects typically have more mass and can generate more energy through nuclear fusion, producing the electromagnetic radiation we observe as luminosity.
	
Additionally, it is worth noting that a galaxy cluster's apparent size and brightness are not always directly proportional to its actual size and luminosity. A cluster's observed size and brightness can also be altered by factors such as its distance from us, the presence of foreground or background objects, and the effects of gravitational lensing. To find the average observed size, let us assume the two galaxy clusters are of equal length. With that, we can emanate how much larger Cluster A is in the sky compared with Cluster B.}

\textit{Demonstrating our assumption may be true, considering that we approximated the distance of Cluster B to us to be $400Mpc$:}
\begin{center}
	\textit{Assume,} $L \propto \mathcal{S}$
	
	$\frac{L_A}{L_B} = \Delta I\frac{d_{A}^2}{d_{B}^2} = 1 
	$
	
	$\boxed{\mathcal{A}_{\mathcal{S}} = \frac{d_B}{d_A} = \frac{50Mpc}{400Mpc} = \frac{1}{8}}
	$
\end{center}
\textit{With this, we deduced that Cluster A amasses eight times more area in the sky than Cluster B. }



\subsection{Calculating the Hubble Constant from redshift and distance}
\textbar{We observe the redshift of Cluster B to be $z_B = 0.1$. With this information, we must relate the redshift to our equations with relativistic redshift to find the Hubble constant.}
\begin{center}
	\textit{Let us define Hubble's constant to be,  }$H_0 = \frac{v}{d}$, and relatistic redshift to be, $z = \frac{v}{c}. $
\end{center}
	\textit{Thus,}
\begin{gather*}
	H_0 = \frac{v}{d} = \frac{cz}{d}
\end{gather*}
\begin{center}
\boxed{H_0 = \frac{cz_B}{d_B} = \frac{(3\times10^6km/s)(0.1)}{400Mpc} = 75kms^{-1}/Mpc}
\end{center}
\textbar{Now we know that the Hubble constant, we can find the redshift of Cluster A. So let us solve for the redshift with our derivation of the relation of Hubbles constant to relativistic redshift.}
\begin{gather*}
	H_0 = \frac{cz}{d} \implies z = \frac{H_0d}{c}
\end{gather*}
\begin{center}
	$\boxed{z_A = \frac{H_0d_A}{c} = \frac{(7.5\times10^4ms^{-1}/Mpc)(50Mpc)}{3\times10^{8}ms^{-1}}=\frac{1}{80}=0.125}
	$
\end{center}

	\section{Special Relativity}
	
\textbar{Special Relativity, or what Einstein initially called the electrodynamics of moving bodies, is a theory that unites electromagnetism and non-accelerating classical mechanical systems. Special Relativity is particularly useful when describing how the motion of objects appears in different reference frames. However, before delving into its derivation, we must be conscious of the two postulates made for Special Relativity. The first postulate is that all physical laws are the same for all inertia frames of reference, regardless of motion. The second postulate is that the speed of light is constant irrespective of where you are in the vacuum of space. }

\subsubsection{Derivation of the Lorentz Transformation in
	Einstein’s Theory of Special Relativity}

	\begin{figure}[h]
	\begin{center}
		
		
		\begin{tikzpicture}
			\draw[very thick,black](-5.5,5) -- (-4.5,5);
			\draw[very thick, purple, -] (-5,0) -- node[left]{$x = \frac{ct}{2}$}   (-5,5);
			\draw[very thick, purple, <->] (-5,1) --   (-5,4);
			\draw[very thick,black](-5.5,0) -- (-4.5,0);
			
		\end{tikzpicture}
	\end{center}
\end{figure}

Suppose we are in a hypothetical spacecraft moving at relativistic speed. Our hypothetical spacecraft contains a small light clock visible from the earth, which aids us in visualizing the differences observed from within and outside the spacecraft. From within the spaceship, we find a light source and two mirrors reflecting the light back and forth. Therefore, according to the observer in the rocket, the light travels in a straight line. 

	\begin{figure}[h]
	\begin{center}	
		\begin{tikzpicture}
			\draw[very thick,black](-4.5,5) -- (-3.5,5);
			\draw[very thick, purple, ->] (-5,0) -- node[left]{}   (-4,5);
			\draw[very thick,black](-5.5,0) -- (-4.5,0);
			\draw[very thick, purple, ->] (-4,5) --  (-3,0);

			\draw[very thick,black](-2.5,0) -- (-3.5,0);
		\end{tikzpicture}	
	\end{center}
\end{figure}

However, from an external observer, let us say we find a contradiction. The light appears not to be travelling in a straight path. With this, we can divide our triangle in two to form a right-angle triangle. We can subsequently use the Pythagorean theorem:

\begin{gather*}
	x^2 =  \left(\frac{ct'}{2}\right)^2 - \left(\frac{vt'}{2}\right)^2
\end{gather*}
\begin{gather*}
x^2 = \frac{c^2{t'}^2}{4} - \frac{v^2{t'}^2}{4}
\end{gather*}
\textit{We know in the other reference frame $x = \frac{ct}{2}$. Thus:}
\begin{gather*}
	\left(\frac{ct}{2}\right)^2 = \frac{c^2{t'}^2 - v^2{t'}^2}{4} 
\end{gather*}
\begin{gather*}
	\frac{c^2t^2}{4} = \frac{{t'}^2(c^2 - v^2)}{4} 
\end{gather*}
\begin{gather*}
	c^2t^2 = {t'}^2(c^2-v^2)
\end{gather*}
\begin{gather*}
	\frac{c^2t^2}{c^2 - v^2} = {t'}^2
\end{gather*}
\begin{gather*}
	{t'}^2 = \frac{t^2}{1 - \frac{v^2}{c^2}}
\end{gather*}
\begin{gather*}
\frac{{t'}}{t} = \sqrt{\frac{1}{1 - \frac{v^2}{c^2}}}
\end{gather*}
\textit{We know $\gamma = \frac{t'}{t}$, thus:}
\begin{center}
	$\boxed{ \gamma = {\frac{1}{\sqrt{1 - \frac{v^2}{c^2}}}}}
	$
\end{center}

\subsubsection{Derivation of relativistic doppler shift}


How does this impact the wavelength of photons emitted by objects at such speeds? For this, we must derive the relativistic doppler shift. 

\textit{Let, $\beta = \frac{v^2}{c^2}$}

\begin{gather*}
\Lambda_{00} = \gamma \implies \Lambda_{0i} = \Lambda_{i0} = -\gamma\beta_i \implies 	\Lambda_{ij} = (\gamma - 1)\left(\frac{\beta_i\beta_j}{\beta^2}\right) + \delta_{ij}
\end{gather*}
\begin{gather*}
\Lambda = \begin{bmatrix} \gamma & -\gamma\beta & 0 & 0 \\
	-\gamma\beta & \gamma & 0 & 0 \\
	0 & 0 & 1 & 0 \\
	0 & 0 & 0 & 1
\end{bmatrix} \implies 	x = \begin{bmatrix} hf_0 \\ -hf_0 \\ 0 \\ 0
\end{bmatrix} x' = \begin{bmatrix} hf_s \\ -hf_s \\ 0 \\ 0 \end{bmatrix}
\end{gather*}
\begin{gather*}
x' = \Lambda x \implies  \begin{bmatrix} hf_s \\ -hf_s \\ 0 \\ 0 \end{bmatrix} = \begin{bmatrix} \gamma & -\gamma\beta & 0 & 0 \\
	-\gamma\beta & \gamma & 0 & 0 \\
	0 & 0 & 1 & 0 \\
	0 & 0 & 0 & 1
\end{bmatrix} \begin{bmatrix} hf_0 \\ -hf_0 \\ 0 \\ 0
\end{bmatrix} 
\end{gather*}
\begin{gather*}
\begin{bmatrix} \gamma h f_0 + \gamma\beta h f_0 \\ -(\gamma\beta hf_0 + \gamma h f_0) \\ 0 \\ 0\end{bmatrix} = \gamma(\beta + 1) \begin{bmatrix} hf_0 \\ -hf_0 \\ 0 \\ 0 \end{bmatrix}
\end{gather*}
\textit{With this, we may rewrite this expression as:}
\begin{gather*}
\frac{1}{f_0} = \gamma\frac{1}{f_s} + \gamma\beta\frac{1}{f_s} \implies \frac{f_s}{f_0} = \gamma + \gamma\beta
\end{gather*}
\begin{gather*}
\frac{f_s}{f_0} = \gamma(1+\beta) = \frac{1}{\sqrt{1-\beta^2}}(1+\beta)
\end{gather*}
\begin{gather*}
	\frac{f_s}{f_0} = \frac{1+\beta}{\sqrt{1-\beta^2}} = \frac{1+\beta}{\sqrt{(1-\beta)(1+\beta)}} = \frac{\sqrt{1+\beta}}{\sqrt{1-\beta}}
\end{gather*}
\begin{center}
	$\boxed{\frac{f_s}{f_0} = \sqrt{\frac{1+\beta}{1-\beta}}}
	$
\end{center}
\textit{We can also correspond this to the wavelength in the relativistic longitudinal doppler effect: }
\begin{center}
	$\boxed{\frac{\Delta\lambda}{\lambda_0} = \sqrt{\frac{1+\beta}{1-\beta}}}
$
\end{center}

To understand these derivations further, let us think of scenarios where we observe objects moving at relativistic speeds. Therefore we shall comprise various situations where this may be useful.

\subsection{Relativistic Doppler Shift Calculation for Two Approaching Spaceships}

Two spaceships (A and B) approach you at 9/10ths the speed of light (0.9c) from opposite directions. They send out radio messages. What speed does one measure for the radio waves from A and B? 


Considering the second postulate of special relativity, we assume the speed of light is constant irrespective of where you are in the vacuum of space. Therefore, since radio waves are electromagnetic radiation, that implies that the radiowaves travels at the speed of light in a vacuum. Therefore, the speed at which we measure the radiowaves emitted by Rocket A and B are the same and are travelling at the speed of light $c$. 


Now what would be the speed of the radio waves from A as measured by B? Since the speed of light is constant for each observer, we may also conclude that the rockets measure each other's radiowaves to be the speed of light $c$.


What is the speed that Spaceship B measures your motion? Considering that each observer is in an inertial frame of reference, each object in this scenario seems to be at rest while the other objects are moving. Therefore, since Spaceship B appears to be approaching the earth at $0.9c$, we know that according to the rocket, it seems the earth is moving towards it at $0.9c$. Thus, Spaceship B measures my motion to be $0.9c$.

\subsubsection{How fast does Spaceship B observe Spaceship A to be moving?}

Utilizing our derivation of the lorentz transformation, we may rewrite the expression in terms of velocity addition: 
\begin{gather*}
\gamma = {\frac{1}{\sqrt{1 - \frac{v^2}{c^2}}}} \implies {v'}_A = \frac{v_A - v_B}{1 - \frac{(v_A)(v_B)}{c^2}} 
\end{gather*}
\begin{center}
	$\boxed{{v'}_A = \frac{v_A - v_B}{1 - \frac{(v_A)(v_B)}{c^2}} =  \frac{0.9c - (-0.9c)}{1 - \frac{(0.9c)(-0.9c)}{c^2}} = 0.9945c}
	$
\end{center}

Therefore, Spaceship B observes Spaceship A to be moving at $0.9945c$.
\subsubsection{Doppler Shift of Radiowaves emitted from relativistic speeds }


Assume the radio messages are emitted from the spaceships with a rest wavelength of 1cm. What is the Doppler Shift that you observe in message A? 

Previously we concluded that Spaceship A was travelling towards us at a speed of 0.9c. With our derivation of the doppler shift, we can find the observed wavelength of the radio wave.
\begin{gather*}
	\frac{\Delta\lambda}{\lambda_0} = \sqrt{\frac{1+\beta}{1-\beta}} \implies 	\Delta\lambda= \lambda_0\sqrt{\frac{1+\beta}{1-\beta}}
\end{gather*}

With this information, we can deduce that the observed wavelength will be the initial wavelength minus the initial wavelength.

\begin{gather*}
	$
	\lambda = \lambda_0 - \Delta\lambda \implies 
$

\end{gather*}


	\section{A Solution to Einsteins Field Equations}
\end{document}
